---
title: "Cog403 Final Project"
author: "Yvonne"
date: '2024-03-25'
output: pdf_document
---

#Installing Packages

```{r, include=FALSE}

#Installing Packages

set.seed(123)

library(tidyr)
library(brms)
library(dplyr)
library(lme4)
library(MASS)
library(kableExtra)
library(tidyverse)
library(broom)
library(nlme)
library(rstanarm)
library(loo)
library(bridgesampling)
library(bayesplot)
library(ggplot2)
```



```{r, include=FALSE}
# loading data

data <- read.csv("SharedResponses_combined_two_genders.csv")
```


```{r}
# getting distinct data points

data <- distinct(data)
```


```{r}
# drop missing data rows

data <- data %>% drop_na()
```


```{r}
# drop certain irrelevant variables

drops <- c("ResponseID", "ExtendedSessionID","Review_political", "UserID")
data <- data[ , !(names(data) %in% drops)]
```

```{r}
# Ensure "Saved" is categorical

data$Saved <- factor(data$Saved)
```

```{r}
# Ensure "Review_poltical_cat" is categorical

data$Review_political_cat <- factor(data$Review_political_cat)
```



```{r}
# Fitting a regular multi linear regression with all variables

model2 <- glm(Saved ~ ., data = data, family = "binomial")
```



```{r}
# Using a stepwise AIC to determine relevant predictors

step <- stepAIC(model2, trace = TRUE, direction= "both")
```

```{r}
# Viewing selected predictors from AIC stepwise funciton

step %>%
  tidy() %>%
  kable(caption = "Regular Logistic Regresion Model")
```




# Bayesian Model

```{r}
# getting priors for non-hierarical bayesian model

get_prior(Saved ~ 1 + PedPed + Barrier + CrossingSignal + AttributeLevel + FemaleExecutive + Review_political_cat, data = data, family = bernoulli())
```


```{r}

# setting priors for a non-hierarical bayesian model

priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("student_t(3,0,2.5)", class = "Intercept")
  )

```


```{r}
#fitting a logistic non-hierarical bayesian regression model (base model)

# Define the model
model_nonhierarchical <- brm(
  formula = Saved ~ 1 + PedPed + Barrier + CrossingSignal + AttributeLevel + FemaleExecutive + Review_political_cat
  , # Model formula
  data = data,
  family = bernoulli("logit"), # Assuming 'Saved' is binary; using logistic regression
  prior = priors,
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 2000,
  seed = 123
)

# Summary of the model
summary(model_nonhierarchical)
```
```{r}
#getting the priors for the logistic hierarical bayesian regression model with politcal orienation as the levels

get_prior(Saved ~ 1 + PedPed + Barrier + CrossingSignal + AttributeLevel + FemaleExecutive + (1 | Review_political_cat), data = data, family = bernoulli())
```



```{r}
# setting the priors for the logistic hierarchical Bayesian model (hierarical model)

priors2 <- c(
  set_prior("normal(0, 2.5)", class = "b"), # For fixed effects
  set_prior("student_t(3,0,2.5)", class = "Intercept"),
  set_prior("student_t(3,0,2.5)", class = "sd") # For random effects standard deviation
)
```


```{r}
# fitting a logistic hierarchical Bayesian model

model_hierarchical <- brm(
  Saved ~ 1 + PedPed + Barrier + CrossingSignal + AttributeLevel + FemaleExecutive + (1 | Review_political_cat),
  data = data,
  family = bernoulli(),
  prior = priors2,
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 2000,
  seed = 123
)
```


```{r}
# visualizing the base model

plot(model_nonhierarchical)
```



```{r}
# visualizing the hierarchical model

plot(model_hierarchical)
```

```{r}
# hierarchical model summary

summary(model_hierarchical)
```

```{r}
# base model summary

summary(model_nonhierarchical)
```


Both models have good convergegence based on rhat values which are close to 1.

The sampling in the non hierarical model may be better (larger sample size) - due to BUlk_ESS and Tail_ESS

WE may prefer the non hierarical model because the results are simler


# Assessing Models

# Trace Plots

```{r}
# hierarical model trace plots

mcmc_trace(model_hierarchical)
```

```{r}
# base model trace plots

mcmc_trace(model_nonhierarchical)
```

# AutoCorrelation Plot
```{r}
# Hierarchical model autocorrelation plots 

mcmc_acf(model_hierarchical)
```

```{r}
# Non hierarical model autocorrelation plots

mcmc_acf(model_nonhierarchical)
```

# R hat (Gelman-Rubin Diagnostic)
```{r}
print(summary(model_hierarchical)$summary[,"Rhat"])
```
```{r}
print(summary(model_nonhierarchical)$summary[,"Rhat"])
```
# Effective Sample Size
```{r}
print(summary(model_hierarchical)$summary[,"n_eff"])
```
```{r}
print(summary(model_nonhierarchical)$summary[,"n_eff"])
```
# Posterior Predicitve check
```{r}
library(bayesplot)

# Set a color scheme that is color-blind friendly
color_scheme_set("blue")

# Perform posterior predictive check with a density overlay and a reasonable number of samples
# Include the observed data (usually the default)
pp_check(model_hierarchical, type = "dens_overlay", nsamples = 50)

# Additionally, customize the plot using ggplot2 functionality
pp_plot <- pp_check(model_hierarchical, type = "dens_overlay", nsamples = 50)

# Use ggplot2 to further customize
library(ggplot2)
pp_plot +
  ggtitle("Posterior Predictive Check: Hierarical Model") +
  xlab("Predicted Values") +
  ylab("Density") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
# Set a color scheme that is color-blind friendly
color_scheme_set("blue")

# Perform posterior predictive check with a density overlay and a reasonable number of samples
# Include the observed data (usually the default)
pp_check(model_nonhierarchical, type = "dens_overlay", nsamples = 50)

# Additionally, customize the plot using ggplot2 functionality
pp_plot <- pp_check(model_hierarchical, type = "dens_overlay", nsamples = 50)

# Use ggplot2 to further customize
pp_plot +
  ggtitle("Posterior Predictive Check: Base Model") +
  xlab("Predicted Values") +
  ylab("Density") +
  theme_minimal() +
  theme(legend.position = "bottom")
```



# Testing the bayesian models


# Leave One Out Cross Validation
```{r}
loo_model_h <- loo(model_hierarchical)
loo_model_nh <- loo(model_nonhierarchical)
```


```{r}
print(loo_model_h)
print(loo_model_nh)
```


```{r}
loo_compare(loo_model_h, loo_model_nh)
```

This measures the predictive accuracy of the models using leave one out cross validation. We can see the hierarchical model has a higher expected log point wise predictive density (ELPD) which indicates it is better at predicting the data. However looking at the standard error of difference between the two models, these results are extremely small and not statistically significant.                      
 

# Watanabe–Akaike Information criterion
```{r}
waic_model_h <- waic(model_hierarchical)
waic_model_nh <- waic(model_nonhierarchical)
```


```{r}
print(waic_model_h)
print(waic_model_nh)
```


# Bayes Factor

```{r}
bridgeh <- bridge_sampler(model_hierarchical)
bridgenh <- bridge_sampler(model_nonhierarchical)
```

```{r}
bf <- bayes_factor(bridgeh, bridgenh)
bf
```

Due to the bayes factor being less than 2, there is evidence in favour of the non hierarical model. It states that the non-hierarical model is more liekly given the data compared to the hierarical one.

```{r}
bf <- bayes_factor(bridgenh, bridgeh)
bf
```


If we take the inverse of the bayes factor we get a value of 30.159 which suggests strong evidence that gien the data the non-hierarchical model is more likely.


# Posterior Predicitve Checks

```{r}
ppd_hierarchical <- posterior_predict(model_hierarchical)
color_scheme_set("brightblue")
pp_check(model_hierarchical, type = "hist", nsamples = 4)


```

```{r}
ppd_combined <- data.frame(iteration = rep(1:4, each = nrow(ppd_hierarchical)),
                           value = c(ppd_hierarchical[,1],
                                     ppd_hierarchical[,2],
                                     ppd_hierarchical[,3],
                                     ppd_hierarchical[,4]))

# Create the histogram with ggplot
ggplot(ppd_combined, aes(x = value, fill = as.factor(iteration))) +
  geom_histogram(alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("#0000FF", "#5555FF", "#AAAAFF", "#FFFFFF")) +
  theme_minimal()
```


```{r}
ppd_nonhierarchical <- posterior_predict(model_nonhierarchical)

color_scheme_set("brightblue")
pp_check(model_nonhierarchical, type = "hist", nsamples = 5)

```



